---
title: "ForeSITE Report Title"
subtitle: "Optional subtitle"
author:
  - name: "Author Name"
    affiliation: "Affiliation"
date: today
format:
  html: default
  typst: default
---

```{r}
#| label: setup
#| include: false
# Brand colors are read from the same _brand.yml that styles this report
brand <- yaml::read_yaml("../_brand.yml")
pal <- unlist(brand$color$palette)
```

## Summary

Replace this text with a short summary of the report. Colors, fonts, and
logos are defined once in `_brand.yml`, so the HTML and PDF versions of this
report stay consistent with ForeSITE pkgdown sites and Shiny apps.

::: {.callout-note}
The ForeSITE crimson (`#A60F2D`) and dark gray (`#4E4E4E`) meet WCAG AA
contrast on white. Use gold (`#FDB921`) only for accents and fills, never for
text on light backgrounds.
:::

## Results

```{r}
#| label: fig-example
#| fig-cap: "Example figure using the ForeSITE palette."
#| fig-alt: "Line chart of two example series over 20 weeks, one in crimson and one in gold."
#| echo: false
set.seed(1)
weeks <- 1:20
series_a <- cumsum(rnorm(20, 1))
series_b <- cumsum(rnorm(20, 0.6))
par(family = "sans", col.axis = pal[["dark-gray"]], col.lab = pal[["dark-gray"]])
plot(
  weeks, series_a, type = "o", pch = 16, lwd = 2, col = pal[["crimson"]],
  ylim = range(series_a, series_b), xlab = "Week", ylab = "Value", bty = "l"
)
lines(weeks, series_b, type = "o", pch = 16, lwd = 2, col = pal[["gold"]])
legend("topleft", c("Series A", "Series B"),
  col = pal[c("crimson", "gold")], lwd = 2, pch = 16, bty = "n"
)
```

@fig-example shows how to reuse the brand palette in base R graphics. For
ggplot2, see `rbranding::brand_set_ggplot()`.

## Acknowledgments {.unnumbered .foresite-acknowledgment}

This report was made possible by cooperative agreement CDC-RFA-FT-23-0069 from
the CDC's Center for Forecasting and Outbreak Analytics. Its contents are
solely the responsibility of the authors and do not necessarily represent the
official views of the Centers for Disease Control and Prevention.
